---
title: "Heterogeneity in susceptibility/exposure"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### HETEROGENEITY IN EXPOSURE 

#### Parameters: 

* `heterogeneity`: determines whether population variation is "**fixed**" (constant over time), "**variable**"  (changes each timestep) or "**homogeneous**" (all units have same potential)
* `correlation`: Is the relationship between transmission & exposure correlated *postively* (pos), *negatively* (neg) or *unrelated* (none)
* `k_transmission`: dispersion parameter for transmission (effective contacts)
* `k_exposure`: dispersion parameter for exposure, required to generate probability distribution when independent of tranmsission
* `R0_estimate`:  mean R0 
* `N` : population size.

Note: if k_exposure = Infinity, this corresponds to the poisson distribution


### POSSIBLE CONDITIONS ##

**Heterogeneity = fixed**
Each unit in the population has a fixed exposure and transmission potential (inherent propert e.g. a static network)

* Correlation = pos (positive)
* Correlation = neg (negative)
* Correlation = none (no correlation)

**Heterogeneity = variable**
The exposure and transmission potential of each unit changes every generation (stochastic property)

* Both exposure potential and transmission potential change each generation, and are not correlated

**Heterogeneity = homogeneous**
Exposure and transmission potential is ocnstant across units and time (perfectly homogeneous)

* Population is perfectly homogeneous with the same exposure and transmission potential


### GENERATION OF EXPOSURE and TRANSMISSION DISTRIBUTIONS ##

```{r, eval = FALSE}


## Condition 1: Heterogeneity = fixed, unit exposure/transmission potential is fixed over generations ##

# Exposure and transmission potential are determined before the generations for-loop 

if (heterogeneity == "fixed"){
  fixed_effc <- rnbinom(N, size = k_transmission, mu = R0_estimate) # generate transmission potential
  
  if(correlation == "pos"){ # if exposure is positively correlated with transmission
    fixed_exp <- fixed_effc # exposure potential == transmission potential 
    
  }else if(correlation == "neg"){ # if exposure is positively correlated with transmission
    fixed_exp <- (1:N)[rev(order(fixed_effc))] # exposure potential is opposite order to transmission potential
    # NB: this is not exact negative correlation, since it orders from numbers 1:N, so units with the same trnasmission potential become ordered with respect to their exposure potential (should have minimal effect with large population)
    
  }else if(correlation == "none"){ # no correlation between exposure and transmission
    fixed_exp <- rnbinom(N, size = k_exposure, mu = R0_estimate) # exposure potential is generated independently with NBD and k_exposure parameter (if set to 1000 = Poisson)
  }
  
}

## Condition 2: Heterogeneity == variable, unit exposure/transmission potential changes every generation ##
# Exposure and transmission potential is re-calculated every generation (code is within the generations for-loop)

if (heterogeneity == "variable"){ # if heterogeneity is variable calculate  exposure and transmission potential independently, no correlation 
  fixed_effc <- rnbinom(N, size = k_transmission, mu = R0_estimate) # tranmsission potential, using k_transmission parameter
  fixed_exp <- rnbinom(N, size = k_exposure, mu = R0_estimate) # exposure potential, usng k_exposure parameter
  
}

## Condition 3: Heterogeneity == homogeneous, unit exposure and transmission potential are equal across time nad across units

if(heterogeneity == "homogeneous"){
  fixed_effc <- rep(R0_estimate, N) # quual transmission potential
  fixed_exp <- rep(R0_estimate, N) # equal probability 
}


```


#### GENERATING CASES WITH HETEROGENEOUS EXPOSURE and TRANSMISSION

* `fixed_effc`: a vector of effective contact potentials of units in the population
+ Generate new I units by subsetting vector for current infected units + summing contact potentials
* `fixed_exp`: a vector of the exposure potential of units in the population
+ Generate new I units by assigning exposure probability to population during sampling


```{r, eval = FALSE}

infected_IDs <- sample(N, I0) # infected IDs in given generation, used to generate new infections


effc <- fixed_effc[infected_IDs] # effc contains effective contact potential of current infected units

# if population is sufficiently large:
effcSum <- sum(effc) # effcSum is the sum of effective contacts for current infected units (an ok assumption when pop is large enough and sampling includes replacement of units)

contact_ids <- unique(sample(N, # population to sample from
                             effcSum, # total effective contacts by all infected units
                             replace=TRUE, # same unit can be contacted by multiple infected units
                             prob = fixed_exp) # probability of a given unit being selected
  )

```

#### Questions:

* Is a negative binomial distribution an appropriate way to model exposure probability?
  + NB: k = 1 is geometric distribution, k = Inf is poisson distribution


```{r testpars}


## Model Definition ##
runs <- 1
generations <- 100 # the number of generations (or timesteps) to run the model over
output <- "counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## PARAMETERS ##

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- 1 # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- 0.1 # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous
k_exposure <- Inf

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?
correlation <- "pos"
if(population_sample == "small"){
  contact_list1 <- replicate(N, c(1:N), simplify = FALSE) 
  contact_list <- lapply(1:length(contact_list1), 
                       function(x) contact_list1[[x]][-x]) 
}else{
  contact_list <- c()
}


## Reseeding Parameters ##
reseed <- FALSE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 0# threshold generation for reseeding events
reseed_frequency <- 0
introductions <- 0 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 0 # start generation for vaccination
V_period <- 0 # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 0 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed

```


```{r testmodel}
Runs_tracker <- state_update(
  ## Population Parameters ##
  state_tracker, # a vector of the the initial state of the population
  N, # population size (may exxclude in favour of calculating in state_update function)
  population_sample, # large or small population (determines sampling strategy)
  
  ## Model Parameters ##
  generations, # the number of generations (or timesteps) to run the model over
  output, # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
  
  ## Epidemiological Parameters ##
  R_period, # the refractory period (number of timesteps that units are immune)
  R0_estimate, # average number of effective contacts per unit
  k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
  k_exposure, # dispersion parameter for exposure, required to generate probability distribution when independent of tranmsission
  heterogeneity, # population variation is “fixed” (constant over time), “variable” (changes each timestep) or “homogeneous” (all units have same potential)
  correlation, # relationship between transmission & exposure correlated postively (pos), negatively (neg) or unrelated (none)
  contact_list, # potential contacts of each unit 
  
  ## Reseeding Parameters ##
  reseed, # reseed 5 infections periodically during the first 10% of generations
  reseed_threshold, # threshold generation for reseeding events
  reseed_frequency,
  introductions, # number of new cases introduced when reseeding occurs
  
  ## Vaccination Parameters ##
  V_prop, # proportion to be vaccinated in each round
  V_start, # start generation for vaccination
  V_period, # duration of immunity from vaccination (by number of timesteps)
  V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
  V_schedule, # how frequent is vacciantion (ie every 2 "years")
  V_mode #random, target or non_target
)

```





