---
title: "The Impact of k on endemicity"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include= FALSE}
library(ggplot2)
library(here)
library(dplyr)
library(reshape2)
library(doParallel)
library(foreach)
library(readr)
#library(RColorBrewer)
library(gridExtra) #for arranging plots side to side with grid.arrange
```

```{r parallelComputing, include = FALSE}

# Set Up for Parallel Computing

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


```


### Functions


```{r functions, include = FALSE}

# Load functions to run model:

source(here("Vfunctions", "V_stateUpdateT2.R")) # wrapper function, tracks epidemic in population over generations
source(here("Vfunctions", "V_DS_T2.R")) # updates state of population based on current generation
source(here("Vfunctions", "S_to_I_T2.R")) # simulation of infection with neg binom dist
source(here("Vfunctions", "effcIdGenT2.R")) # generating effective contact IDs
source(here("Vfunctions", "effcChecker.R")) # verifying effective contact IDs
source(here("Vfunctions", "reseedingT2.R")) # reseeding infection

```

Include function for calculating median incidence for last 100 generations:

```{r endStatsFn, include = FALSE}

# Code to calculate median incidence at equilibrium (after 900Gens)

endStats <- function(mat){
  eq <- mat[, (ncol(mat)-100) :ncol(mat)] # subset last 100 generations of matrix
  id <- which(eq[,ncol(eq)] != 0) # identify which simulations become endemic (cases > 0 in final generation)
  if(length(id) == 0){ # if no simulations become endemic then all statistics set to 0
    M <- 0 
    iqr <- 0
    lq <- 0
    uq <- 0
    
  }else if( length(id) > 0 ){ # for simulations which become endemic...
    mEnd <- as.vector(eq[id,]) # subset last 100 generations of endemic simulations and format as vector
    # take the median, iqr, lower and upper quartiles over the last 100 generations for all simulations which become endemic 
    M <- median(mEnd) 
    iqr <- IQR(mEnd)
    lq <- quantile(mEnd)[2] # lower quartile
    uq <- quantile(mEnd)[4] # upper quartile
  }
  sumStat <- c(Med = M, iqr = iqr, lq = lq, uq = uq) # concatenate statistics
  return(sumStat)  # return vector of summary stats
  }
```

Include function for calculating the proportion of simulations which become endemic for a given parameter set.
```{r pEndemicity, include = FALSE}

# Calculate proportion of simulations which become endemic

endemicity <- function(Tmat){ # input function is matrix of numbers infected across generations (columns) for multiple simulations (rows)
  
  P_endemic <- sum(Tmat[,ncol(Tmat)]!=0) / nrow(Tmat) # sum the number of simulations which result in 0 infections in the final generation (extinction), divided by the total number of simulaitons (number of row of matrix)

  return(P_endemic) # return the proportion of simulations which become endemic
}

```


### What is the impact of transmission heterogeneity on endemicity?

#### Model Parameters:

* Explore variable levels of heterogeneity from k = 0.1 to k = Inf
  + *see negative binomial distribution plot for explanation of k parameter*
* Explore model behaviour over a range of values for R0 and refractory periods to understand how the parameters interact.
* Include reseeding to prevent early stochastic die-out
* Run for 1000 generations to observe long term trajectory

#### The k parameter

* The negative binomial distribution is used to describe the distribution of $v$ the individual reproduction number (effective contact rate of units)
* Following Lloyd-Smith *et al.* (2005), $v$ is gamma-distributed with mean R0 and dispersion parameter k. The number of secondary infections caused by each case, $Z$ is described by offxpring distribution $Z∼ negativebinomial(R0,k)$ 

* $R0$ is defined as the mean number of infections caused by a signle infected individual in an entirely susceptible population 
* $k$ provides a measure of dispersion around the mean $R0$, where smaller values of $k$ indicate greater heterogeneity (Lloyd-Smith *et al.* 2005)
  + The Poisson ($k\to\infty$) and geometric ($k = 1$) models are special cases of the negative binomial distribution.
  + $k\to\infty$ : in generation-based models neglecting individual variation, $v = R0$ for all cases, yielding $Z∼Poisson(R0)$
  + $k = 1$ : in differential-equation models with homogeneous transmission and constant recovery rates, $v$ is exponentially distributed, yielding $Z∼geometric(R0)$
* The plot below shows the proportion of cases accounted for by the top 10% of infectious units under different values of $k$. 


```{r NBDcode, include = FALSE}

nbd_k <- c(seq(0.1, 1, 0.1))
nbd_k2 <- c(1:10, Inf)
nbd_R0 <- c(1,3,5)

nbd_pars <- expand.grid(nbd_k, nbd_R0)
colnames(nbd_pars) <- c("k", "R0")
nbd_pars$prop <- 0

nbd_pars2 <- expand.grid(nbd_k2,nbd_R0)
colnames(nbd_pars2) <- c("k", "R0")
nbd_pars2$prop <- 0

N <- 1e5
p <- 0.1 #top 10% of infectious individuals
x <- 0.1*N

for(i in 1:nrow(nbd_pars)){
  v <- rnbinom(N, mu = nbd_pars[i,2], size = nbd_pars[i,1])
  
  nbd_pars[i,3] <- sum(rev(sort(v))[1:x]) / sum(v)
}  


for(i in 1:nrow(nbd_pars2)){
  v <- rnbinom(N, mu = nbd_pars2[i,2], size = nbd_pars2[i,1])
  
  nbd_pars2[i,3] <- sum(rev(sort(v))[1:x]) / sum(v)
} 

nbd_pars2[,"k"] <- ordered(nbd_pars2[,"k" ])

```

```{r NBDplots, echo=FALSE}
NBD1 <- ggplot(data = nbd_pars, aes(x = factor(k), y = prop, group = R0, colour = factor(R0)))+
  geom_line()+
  xlab("k value")+
  ylab("Proportion of transmission by top 10% infectious units")+
  scale_color_discrete("R0")+
  scale_y_continuous(limits = c(0,1))+
  theme_bw() 

NBD2 <- ggplot(data = nbd_pars2, aes(x = k, y = prop, group = R0, colour = factor(R0)))+
  geom_line()+
  xlab("k value")+
  ylab("Proportion of transmission by top 10% infectious units")+
  scale_color_discrete("R0")+
  scale_y_continuous(limits = c(0,1))+
  theme_bw()

grid.arrange(NBD1, NBD2, ncol = 2)

```


```{r data, include = FALSE}

## LOAD IN THE DATA ##

kpars <- c(0.1,0.3,0.5,0.7,1,Inf)

load(here("data_080121/k0-1List.RData"))
load(here("data_080121/k0-3List.RData"))
load(here("data_080121/k0-5List.RData"))
load(here("data_080121/k0-7List.RData"))
load(here("data_080121/k1-List.RData"))
load(here("data_080121/kInfList.RData"))

masterList <- c(k0.1List,
                k0.3List, 
                k0.5List, 
                k0.7List, 
                k1List, 
                kInfList)

R0Rp <- read.csv(here("data_080121/pars.csv"))

R0Rp<- (rbind(R0Rp,R0Rp,R0Rp,R0Rp,R0Rp,R0Rp))

pars <- cbind("k" = rep(kpars,each = 60), R0Rp)

```

### The Impact of Heterogeneity on Incidence and Endemicity:

#### Plot 1 : Median Incidence at Equilibrium

A plot which shows the median incidence and IQR (from generation 900-1000) for different values of R0 (x axis), Refractory Period (facets), and k parameters (lines). 

```{r Incidence, echo = FALSE}

# Calculating summary statistics for incidence at equilibrium:

EqStats <- t(sapply(masterList, endStats)) # applying endStats function to data
medIncidence <- as.data.frame(cbind(pars, EqStats)) # creating a dataframe with summary stats of incidence at equilibrium with parameter scenarios.

## Plot of median incidence (and IQR) at equilibrium with increasing R0 (x axis)  and refractory periods as separate facets ##
# NOTE: Where a proportion of simulations run to extinction, the median incidence + IQR only includes simulations which result in endemicity.

# The plots are qualitatively very similar, below is a plot with the extreme values of k (0.1, Inf) to demonstrate this.
Incidenceplot <- medIncidence %>% filter(k == 0.1 | k == Inf)
ggplot(data = Incidenceplot, aes(x = R0, y = Med, color = factor(k)))+
  geom_line()+
  #geom_point()+
  geom_errorbar(aes(ymin=lq, ymax=uq), 
                width = 0.3, # specifies width of whiskers (max/min)
                position = position_dodge(width = 0.1) # specifies spacing between lines
                )+
  facet_wrap(~Rperiod, labeller = "label_both")+
  scale_x_continuous(breaks = seq(0,5,1))+
  xlab("R0")+
  ylab("Median incidence at Equilibrium")+
  labs(colour = "k value")+
  theme_bw()

```

**What is the impact of $k$ on the median incidence at equilibrium (after 900generations)?**

* For most $R0$ and refractory period combinations, the trend in median incidence observed is robust to changes in the value of $k$ (level of heterogeneity), with only small differences observed between the most extreme k values ( $k = 0.1; k=\infty$)
  + For refractory periods > 5 generations, extinction occurs at lower values of $R0$ when $k = 0.1$ (high heterogeneity) compared to $k = \infty$ (homogeneity)
* For values of $R0$ lower than the threshold $R0$ (corresponding to maximum incidence), variation around the median incidence is greater when $k = 0.1$ (high heterogeneity) compared to $k = \infty$ (homogeneity).
  + At values of $R0$ above the threshold the variation around the median incidence is similar when $k = 0.1$ (high heterogeneity) and $k = \infty$ (homogeneity).  
  
  
```{r oscillations, eval = FALSE, include = FALSE}

## Model Definition ##
runs <- 1
generations <- 1000 # the number of generations (or timesteps) to run the model over
output <- "I_counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- 3 # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- Inf # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?
contact_list1 <- replicate(N, c(1:N), simplify = FALSE) 
contact_list <- lapply(1:length(contact_list1), 
                       function(x) contact_list1[[x]][-x]) 
if(population_sample == "large"){
  contact_list <- 0
}


## Reseeding Parameters ##
reseed <- TRUE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 100 # threshold generation for reseeding events
reseed_frequency <- 10
introductions <- 10 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 0 # start generation for vaccination
V_period <- 0 # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 0 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed



## MODEL ## 
k_vals <- c(0.1, Inf)
generations <- 10000
trajectory_df <- vector()

for(i in 1:length(k_vals)){
  k_transmission <- k_vals[i]
  
  Runs_tracker <- 
      #foreach (z = 1:runs, .combine = "rbind") %dopar% {
       state_update(
         
          state_tracker, # a vector of the the initial state of the population
          N, # population size (may exxclude in favour of calculating in state_update function)
          population_sample, # large or small population (determines sampling strategy)
          
          generations, # the number of generations (or timesteps) to run the model over
          output, # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
          
          R_period, # the refractory period (number of timesteps that units are immune)
          R0_estimate, # average number of effective contacts per unit
          k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
          heterogeneity, # is heterogeneity in effective contacts "fixed" or "variable"
          contact_list, # potential contacts of each unit 
          
          reseed, # reseed 5 infections periodically during the first 10% of generations
          reseed_threshold, # threshold generation for reseeding events
          reseed_frequency,
          introductions, # number of new cases introduced when reseeding occurs
          
          V_prop, # proportion to be vaccinated in each round
          V_start, # start generation for vaccination
          V_period, # duration of immunity from vaccination (by number of timesteps)
          V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
          V_schedule, # how frequent is vacciantion (ie every 2 "years")
          V_mode #random, target or non_target
        )
  
  trajectory <- c("k" = k_vals[i], Runs_tracker)
  trajectory_df <- as.data.frame(rbind(trajectory_df, trajectory))
  
}

write.csv(trajectory_df, file = "ktrajectory.csv", row.names = FALSE)
  
```

**Why is there greater variation for lower k values?**

* As seen in the plot below, when the population has a high level of heterogeneity (k = 0.1) there is greater variation around the median incidence, a pattern which is conserved long-term (10,000 generations). 
* The high level of heterogeneity between units in terms of transmission potential explains this pattern. Stochasticity in the infection process means that at some time points the majority of new infections may be high-contact units - driving a rapid increase in incidence, whilst at other timepoints the majority of new infections may be units with few, or no, contacts - resulting in a decline to very low incidence.
* The lower variation around median incidence in the k = Inf scenario, reflects the increased homogeneity of the population, with less variation in unit contact rates resulting in oscillation over a narrow range of incidence.

```{r oscillationPlot, echo = FALSE, message = FALSE, warning = FALSE}

trajectory_df <- read_csv(here("data_080121/ktrajectory.csv"))
trajectory_dfLong <- melt(trajectory_df, id.vars = "k")

ggplot(data = trajectory_dfLong, aes(x = variable, y = value))+
  geom_line(aes(group = 1))+
  facet_wrap(~k)+
  scale_x_discrete(breaks = seq(0,10000,1000), labels = as.character(seq(0,10000,1000)))+
  xlab("Generation")+
  ylab("Incidence")+
  theme_bw()

```

#### Plot 2 : Proportion of simulations which become endemic

The plot shows how different values of Refractory Period (x axis), $R0$ (facets) and $k$ (lines) impact the proportion of simulations resulting in endemicity.

```{r endemicity, echo = FALSE}


pEndemicity <- sapply(masterList, endemicity) # Apply endemicity function to kInfList to calculate proportion of simulations which become endemic for different parameter combinations
  
pEndemicitydf <- as.data.frame(cbind(pars,pEndemicity)) # Create a dataframe of proportion endmeic simulations and respective parameter combinations

## Plot of proportion of simulations resulting in endemicity, with increasing refractory periods, and R0 values shown in separate facets ##

ggplot(data = pEndemicitydf, aes(x = Rperiod, 
                                 y = pEndemicity,
                                 color = factor(k)
                                 ))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks = seq(1,11,2))+
  xlab("Refractory Period")+
  ylab("Proportion Endemicity (100 sims)")+
  facet_wrap(~R0, labeller = "label_both", scales = "free_x")+
  scale_colour_discrete("k Value")+
  theme_bw()

```

**What is the impacts of $k$ on the proportion of simulations which result in extinction?**  

For $R0 < 1$ : 

* All simulations result in extinction

Between $R0 = 1.5$ and $R0 = 3.5$

* For a given $R0$ and refractory period combination, as the value of $k$ decreases (population becomes more heterogeneous), the proportion of simulations which result in extinction increases. 
* For a given $R0$, the threshold beyond which a proportion of simulations result in extinction occurs at shorter refractory periods as $k$  decreases (population heterogeneity increases).

For $R0 > 4$

* The trend in endemicity as the refractory period increases is unaffected by changes in the $k$ parameter (population heterogeneity)

In general: 

* For a given set of parameters ($R0$, Refractory Period), the likelihood of extinction increases as population heterogeneity increases.
* From Lloyd-Smith *et al.* (2005): 
  + *"Extinction risk rises owing to a higher proportion of non-transmitting cases when $v$ (individual reproduction number) is overdispersed"*
  + *"This effect thwarts invasion by diseases that are very potent spreaders on average: for arbitrarily high R0* $q \to\ 1$ (q = extinction probability) *as* $k\to\0$"

**Why does the probability of extinction increase with decreasing k value?**

* For the parameter set: $R0 = 2, R\_period = 7$ the plots below the epidemic trajectory of multiple simulations for different k values.
* The plot shows that for lower values of k there is greater variation in incidence, with very high and low incidences apparent at different timepoints (see above for explanation). 
* For all k parameter scenarios the epidemic is oscillatory, with rapid increases (epidemic peak) and decreases (epidemic troughs) in incidence. 
  + During epidemic troughs, extinction is possible due to the sequential infection of low(/zero)-contact individuals (in this case assuming depletion of susceptibles is not the limiting factor- verified 150221).
  + In a heterogeneous population, there is a greater proportion of low/no-contact units and thus the probability of stochastic extinction is increased. 

```{r kEndemicityTrajectory, eval = FALSE, include = FALSE}

R0_estimate <- 2
R_period <- 7
k_vals <- c(0.1, 0.3, Inf)
runs <- 5

## Model 


EndTrajList <- list()

for(i in 1:length(k_vals)){
    
    k_transmission <- k_vals[i]
    
    print(i)
    
    Runs_tracker <-
      foreach (z = 1:runs, .combine = "rbind") %dopar% {
       state_update(
         
          state_tracker, # a vector of the the initial state of the population
          N, # population size (may exxclude in favour of calculating in state_update function)
          population_sample, # large or small population (determines sampling strategy)
          
          generations, # the number of generations (or timesteps) to run the model over
          output, # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
          
          R_period, # the refractory period (number of timesteps that units are immune)
          R0_estimate, # average number of effective contacts per unit
          k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
          heterogeneity, # is heterogeneity in effective contacts "fixed" or "variable"
          contact_list, # potential contacts of each unit 
          
          reseed, # reseed 5 infections periodically during the first 10% of generations
          reseed_threshold, # threshold generation for reseeding events
          reseed_frequency,
          introductions, # number of new cases introduced when reseeding occurs
          
          V_prop, # proportion to be vaccinated in each round
          V_start, # start generation for vaccination
          V_period, # duration of immunity from vaccination (by number of timesteps)
          V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
          V_schedule, # how frequent is vacciantion (ie every 2 "years")
          V_mode #random, target or non_target
        )
        
      }
    
    
    EndTrajList[[i]] <- Runs_tracker
    
  }



EndTrajdf <- as.data.frame(do.call("rbind", EndTrajList))
colnames(EndTrajdf) <- 1:1000
EndTrajdf$run <- rep(1:5, 3)
EndTrajdf$k <- rep(c(0.1,0.3,Inf), each = 5)
write.csv(EndTrajdf, file = "kEndemicityTrajectory.csv", row.names = FALSE)


```

```{r kEndTrajPlot, echo = FALSE, message = FALSE, warning = FALSE}

EndTrajdf <- read_csv(here("data_080121/kEndemicityTrajectory.csv"))

longEndtraj <- melt(EndTrajdf, id.vars = c("k", "run"))

ggplot(data = longEndtraj, aes(x = variable, y = value))+
  geom_line(aes(group = 1), size = 0.1)+
  xlab("Generation")+
  ylab("Incidence")+
  scale_x_discrete(breaks = seq(0,1000,200), labels = as.character(seq(0,1000,200)))+
  facet_grid(k~run, labeller = label_both)+
  theme_bw()

```


```{r, include = FALSE, eval = FALSE}


## **Alternative Plot** ##

#The plot shows how endemicity is impacted by $R0$ (x axis), Refractory Period (facets) and $k$

ggplot(data = pEndemicitydf, aes(x = R0, y = pEndemicity, color = factor(k)))+
  geom_line()+
  geom_point()+
  facet_wrap(~Rperiod, labeller = "label_both")+
  xlab("R0")+
  ylab("Proportion Endemicity (100 sims)")+
  scale_colour_discrete("k")+
  theme_bw()

```