---
title: "R0 Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r libraries, include= FALSE}
library(ggplot2)
library(here)
library(dplyr)
library(reshape2)
library(doParallel)
library(foreach)

```

```{r parallelComputing, include = FALSE}

# Set Up for Parallel Computing

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


```


### Functions

Load functions to run model, with most recent version

```{r functions}
source(here("Vfunctions", "V_stateUpdateT2.R")) # wrapper function, tracks epidemic in population over generations
source(here("Vfunctions", "V_DS_T2.R")) # updates state of population based on current generation
source(here("Vfunctions", "S_to_I_T2.R")) # infection simulation
source(here("Vfunctions", "effcIdGenT2.R")) # generating effective contact IDs
source(here("Vfunctions", "effcChecker.R")) # verifying effective contact IDs
source(here("Vfunctions", "reseedingT2.R")) # reseeding infection

```


### Test 1 : A single epidemic

Simulating a single epidemic:

* Runs = 1, for time saving
* No long-lasting immunity: $R\_period = 1$
* Ignore: Reseeding (not simulating endemicity), Vaccination (only want to explore epidemic)



```{r pars1}

## Model Definition ##
runs <- 1
generations <- 100 # the number of generations (or timesteps) to run the model over
output <- "counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## PARAMETERS ##

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- Inf # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- Inf # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?
contact_list1 <- replicate(N, c(1:N), simplify = FALSE) 
contact_list <- lapply(1:length(contact_list1), 
                       function(x) contact_list1[[x]][-x]) 

## Reseeding Parameters ##
reseed <- FALSE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 0# threshold generation for reseeding events
reseed_frequency <- 0
introductions <- 0 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 0 # start generation for vaccination
V_period <- 0 # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 0 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed

```

#### Run Model: 

```{r sim1}
Runs_tracker <-
  state_update(
    state_tracker, 
    N, 
    population_sample,
    generations,
    output, 
    R_period, 
    R0_estimate,
    k_transmission, 
    heterogeneity,
    contact_list,
    reseed,
    reseed_threshold,
    reseed_frequency,
    introductions,
    V_prop, 
    V_start,
    V_period,
    V_rounds,
    V_schedule,
    V_mode
  )

```

#### Plot Model:

```{r plot1}

    Runs_tracker$state <- c("S","I","R","V")
    Runs_long <- melt(Runs_tracker, id.var = "state")
    
    ggplot(data = Runs_long, aes(x = variable,y=value, group = state))+
      geom_line(aes(colour = state))+
      scale_x_discrete(breaks = seq(0,100,10))+
      xlab("Generation")+
      ylab("Population")+
      scale_fill_discrete("Disease State")+
      theme_bw()
 
```

### What is the Impact of R0?

* How does R0 impact 
  (i) the timing of the epidemic (plot epidemic trajectories overlaid on top of one another)
  (ii) the magnitude of the epidemic (plot the cumulative incidence)
  
  
Parameters:
* Set R0 values between 0 - 5 to explore behaviour around R0 = 1 and as R0 increases

```{r R0impact, eval = FALSE}

output <- "I_counts" # in order to compare epidemic trajectory with increasing values of R0
R0_vals <- seq(0,5,0.5)
runs <- 100

runs_list <- list()

system.time(
for(r in 1:length(R0_vals)){
  
  R0_estimate <- R0_vals[r]
  
  Runs_tracker <-
    foreach (z = 1:runs, .combine = "rbind") %dopar% {  
      state_update(
        state_tracker, 
        N, 
        population_sample,
        generations,
        output, 
        R_period, 
        R0_estimate,
        k_transmission, 
        heterogeneity,
        contact_list,
        reseed,
        reseed_threshold,
        reseed_frequency,
        introductions,
        V_prop, 
        V_start,
        V_period,
        V_rounds,
        V_schedule,
        V_mode
      )
      
      #runs_ <- rbind(runs_df, Runs_tracker)
      
    }
  runs_list[[r]] <- Runs_tracker
}
)



```

Plot Epidemic trajectories with different values of R0 

(i) Impact of increasing R0 on timing of epidemic in a closed population with lifelong immunity
* As the value of R0 increases, the timing of the epidemic is affected in two ways: (i) the epidemic peak occurs sooner when R0 is bigger and (ii) the epidemic period is shorter with a steep incline in cases followed by a sharp decline to extinction. 

(ii) Impact of increasing R0 on **magnitude** of epidemic in a closed population with lifelong immunity
* As the value of R0 increases from 0 - 5 the magnitude of the epidemic increases. At values of $R0 < 1.5$ the outbreak rapidly declines to extinction with no secondary spread from the index cases. As the value of R0 increases from 1.5, the magnitude of the outbreak also increases. At $R0 = 1.5$ the number of cases at the epidemic peak is 628 (6.28% of the population), at $R0 = 5$ the number of cases at the peak of the epidemic is 4530 (45.3% of the population).
* The threshold effect at $R0 = 1$ is also seen in the scatterplot, which shows a sigmoid curve, with a steep increase in cumulative incidence between $R0 = 1$ and $R0 = 1.5$. At $R0 = 1$ there is variation about the mean, reflecting that in some cases there will be some disease spread before extinction, whilst in others the epidemic will decline to extinction immediately. 
* These results correspond to the expected behaviour around the epidemic threshold of R0 = 1

```{r R0plots}

load(here("data_080121/WU1/R0singleEpidemic_list.RData"))
runs_list <- R0list_singleEpidemic
R0_vals <- R0_vals <- seq(0,5,0.5)

meanEpidemic <- as.data.frame(t(sapply(runs_list, 
                       function(x) apply(x, 2, mean)
                       )))

#rownames(meanEpidemic) <- R0_vals
meanEpidemic$R0 <- R0_vals
meanEpidemic_long <- melt(meanEpidemic, id.vars = "R0")
colnames(meanEpidemic_long) <- c("R0", "gen", "cases")
    
ggplot(data = meanEpidemic_long, aes(x = gen,y=cases, group = factor(R0)))+
      geom_line(aes(colour = factor(R0)))+
      scale_x_discrete(breaks = seq(0,100,10))+
      xlab("Generation")+
      ylab("Cases")+
      scale_colour_discrete("R0 values")+
      theme_bw()

#Cumualtive Incidence

meanCI <- sapply(runs_list, function(x)
  mean(apply(x, 1, sum)))

sdCI <- sapply(runs_list, function(x)
  sd(apply(x, 1, sum)))


CumInc <-as.data.frame(cbind(R0_vals, 
                             "meanCI" =meanCI, 
                             "sdCI" =sdCI))

ggplot(data = CumInc, aes(x = R0_vals,y=meanCI))+
  geom_point()+
  geom_errorbar(aes(ymin=meanCI-sdCI, ymax=meanCI+sdCI), width=.2)+
  xlab("R0")+
  ylab("Cumulative Incidence")+
  theme_bw()
```

