---
title: "Refractory Period Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r libraries, include= FALSE}
library(ggplot2)
library(here)
library(dplyr)
library(reshape2)
library(doParallel)
library(foreach)

```

```{r parallelComputing, include = FALSE}

# Set Up for Parallel Computing

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


```

### Functions

Load functions to run model, with most recent version

```{r functions}
source(here("Vfunctions", "V_stateUpdateT2.R")) # wrapper function, tracks epidemic in population over generations
source(here("Vfunctions", "V_DS_T2.R")) # updates state of population based on current generation
source(here("Vfunctions", "S_to_I_T2.R")) # infection simulation
source(here("Vfunctions", "effcIdGenT2.R")) # generating effective contact IDs
source(here("Vfunctions", "effcChecker.R")) # verifying effective contact IDs
source(here("Vfunctions", "reseedingT2.R")) # reseeding infection

```

```{r usdfunction}

# function to calculate median incidence at IQR for last 100 generations
EpiStat <- function(mat) {
  if (median(mat) == 0) {
    M <- median(mat)
    iqr <- IQR(mat)
    q1 <- quantile(mat)[2]
    q3 <- quantile(mat)[4]
  } else{
    vec <- as.vector(mat[(mat != 0)])
    M <- median(vec)
    iqr <- IQR(vec)
    q1 <- quantile(vec)[2]
    q3 <- quantile(vec)[4]
  }
  sumStat <- c(M, iqr, q1, q3)
  return(sumStat)
}

```


### What is the impact of the refractory period?

* $generations = 1000$ to view long term trajectory
* $R0 = 3$ R0 > 1 to explore impact on disease transmission & epidemic trajectory/endemicity
* Refractory periods between 1 and 100 to explore impact of short and long term immunity

```{r pars1, eval = FALSE}

## Model Definition ##
runs <- 1
generations <- 100 # the number of generations (or timesteps) to run the model over
output <- "counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## PARAMETERS ##

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- Inf # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- Inf # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?
contact_list1 <- replicate(N, c(1:N), simplify = FALSE) 
contact_list <- lapply(1:length(contact_list1), 
                       function(x) contact_list1[[x]][-x]) 

## Reseeding Parameters ##
reseed <- FALSE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 0# threshold generation for reseeding events
reseed_frequency <- 0
introductions <- 0 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 0 # start generation for vaccination
V_period <- 0 # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 0 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed

```


```{r RpImpact, eval = FALSE}

generations <- 1000
R0_estimate <- 3
Rp_vals <- seq(1,10,1)

Rp_list <- list()

system.time(
for(r in 1:length(Rp_vals)){
  
  R_period <- Rp_vals[r]
  
  Runs_tracker <-
    foreach (z = 1:runs, .combine = "rbind") %dopar% {  
      state_update(
        state_tracker, 
        N, 
        population_sample,
        generations,
        output, 
        R_period, 
        R0_estimate,
        k_transmission, 
        heterogeneity,
        contact_list,
        reseed,
        reseed_threshold,
        reseed_frequency,
        introductions,
        V_prop, 
        V_start,
        V_period,
        V_rounds,
        V_schedule,
        V_mode
      )
      
    }
  Rp_list[[r]] <- Runs_tracker
}
)

```

Plot epidemic trajectories with different refractory periods

**(i) Impact on Endemicity**

* As the refractory period increases, the outbreak more often results in extinction. With $R0 = 3$ and a homogeneous population, lower refractory periods (of 1 and 5 generations) result in disease endemicity, whilst higher refractory periods (of 10-100) result in extinction. A longer refractory period leads to more units remaining immune to infection at a specific time; at a given threshold refractory period, the susceptible population will be depleted such that infection can no longer be sustained and therefore declines to extinction. 

**(ii) Impact on average incidence at equilibrium (see plot 2)**

* As the refractory period increases, the average incidence at equilibrium decreases (to extinction for long refractory periods). For outbreaks which become endemic (refractory periods of 1 and 5), an increase in the length of the refractory period corresponds to greater variation in incidence, and a longer delay before equilibrium is reached. 

```{r rpPlots}

## Mean Epidemic Trajectory ##

load(here("data_080121", "WU1", "rpImpact.RData"))
Rp_vals <- seq(1,10,1)


rpEpidemic <- t(sapply(Rp_list, 
                       function(x) apply(x, 2, mean)
                       ))

rownames(rpEpidemic) <- Rp_vals
rpEpidemic_long <- melt(rpEpidemic)
colnames(rpEpidemic_long) <- c("R_period", "gen", "cases")
rpEpidemic_long <- rpEpidemic_long %>% filter(R_period == 1 |
                                                  R_period == 3 |
                                                  R_period == 5 |
                                                  R_period == 7 |
                                                  R_period == 9 )
    
ggplot(data = rpEpidemic_long, aes(x = gen,y=cases, group = factor(R_period)))+
      geom_line(aes(colour = factor(R_period)))+
      xlab("Generation")+
      ylab("Cases")+
  scale_colour_discrete("Refractory Period")+
      theme_bw()
     # coord_cartesian(xlim = c(0,250))

## Incidence at Equilibrium ##

Rp_Eq <- lapply(Rp_list, 
                function(x) 
                  x[, 900:1000])

sumStat <- t(sapply(Rp_Eq, EpiStat))
colnames(sumStat) <- c("med", "iqr", "q1", "q3")
Rp_CumInc <- as.data.frame(cbind("Rperiod" =Rp_vals, sumStat))

# plot of median incidence
ggplot(data = Rp_CumInc, aes(x = factor(Rp_vals),y=med))+
  geom_point()+
  geom_errorbar(aes(ymin=q1, ymax=q3), width=.2)+
  xlab("Refractory Period")+
  ylab("Incidence at Equilibrium (G900)")+
  theme_bw()

```

