---
title: "Vaccination and Heterogeneity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include= FALSE}
library(ggplot2)
library(here)
library(dplyr)
library(reshape2)
library(doParallel)
library(foreach)
library(readr)
#library(RColorBrewer)
#library(gridExtra) - for arranging plots side to side with grid.arrange
```

```{r parallelComputing, include = FALSE}

# Set Up for Parallel Computing

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


```


```{r functions, include = FALSE}

### Functions

# Load functions to run model:

source(here("addingExposure/vSU_exp.R")) # wrapper function, tracks epidemic in population over generations
source(here("addingExposure/vDS_exp.R")) # updates state of population based on current generation
source(here("addingExposure/vSI_exp.R")) # simulation of infection with neg binom dist
source(here("addingExposure/veffcIdGen_exp.R")) # generating effective contact IDs
source(here("Vfunctions", "effcChecker.R")) # verifying effective contact IDs
source(here("Vfunctions", "reseedingT2.R")) # reseeding infection

```


```{r endStatsFn, include = FALSE}

#Include code for calculating median incidence for last 100 generations:

# Code to calculate median incidence at equilibrium (after 900Gens)

endStats <- function(mat){
  eq <- mat[, (ncol(mat)-100) :ncol(mat)] # subset last 100 generations of matrix
  id <- which(eq[,ncol(eq)] != 0) # identify which simulations become endemic (cases > 0 in final generation)
  if(length(id) == 0){ # if no simulations become endemic then all statistics set to 0
    M <- 0 
    iqr <- 0
    lq <- 0
    uq <- 0
    
  }else if( length(id) > 0 ){ # for simulations which become endemic...
    mEnd <- as.vector(eq[id,]) # subset last 100 generations of endemic simulations and format as vector
    # take the median, iqr, lower and upper quartiles over the last 100 generations for all simulations which become endemic 
    M <- median(mEnd) 
    iqr <- IQR(mEnd)
    lq <- quantile(mEnd)[2] # lower quartile
    uq <- quantile(mEnd)[4] # upper quartile
  }
  sumStat <- c(Med = M, iqr = iqr, lq = lq, uq = uq) # concatenate statistics
  return(sumStat)  # return vector of summary stats
  }
```

```{r pEndemicity, include = FALSE}
#Include code for calculating the proportion of simulations which become endemic for a given parameter set.


# Calculate proportion of simulations which become endemic

endemicity <- function(Tmat){ # input function is matrix of numbers infected across generations (columns) for multiple simulations (rows)
  
  P_endemic <- sum(Tmat[,ncol(Tmat)]!=0) / nrow(Tmat) # sum the number of simulations which result in 0 infections in the final generation (extinction), divided by the total number of simulaitons (number of row of matrix)

  return(P_endemic) # return the proportion of simulations which become endemic
}

```


### How does heterogeneity in Exposure and Transmission impact endemicity


Check behaviour with following parameter combinations:

**Try with greater range of k parameters, and/or {R0:Rperiod} scenarios

* Heterogeneity == Fixed (to observe impact of correlation between exp/transmission)
* Correlation: positive, negative, none
* Dispersion: 0.1, Inf

Positive correlation:

* k_transmission = 0.1, k_exposure = 0.1
* k_transmission = Inf, k_exposure = Inf

Negative correlation:

* k_transmission = 0.1, k_exposure = 0.1 (but opposite order of units)
* k_transmission = Inf, k_exposure = Inf

None: 

* k_transmission = 0.1, k_exposure = 0.1 
* k_transmission = Inf, k_exposure = Inf
* k_transmission = 0.1, k_exposure = Inf



```{r testpars, eval = FALSE, include = FALSE}


## Model Definition ##
runs <- 100
generations <- 1000 # the number of generations (or timesteps) to run the model over
output <- "I_counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## PARAMETERS ##

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- 1 # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- 0.1 # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous
k_exposure <- Inf

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?
correlation <- "pos"
if(population_sample == "small"){
  contact_list1 <- replicate(N, c(1:N), simplify = FALSE) 
  contact_list <- lapply(1:length(contact_list1), 
                       function(x) contact_list1[[x]][-x]) 
}else{
  contact_list <- c()
}


## Reseeding Parameters ##
reseed <- TRUE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 100 # threshold generation for reseeding events
reseed_frequency <- 10
introductions <- 10 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 850 # start generation for vaccination
V_period <- Inf # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 0 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed

```



```{r pars, include = FALSE}


Correlation <- rep(c(rep("pos", 2), rep("neg",2), rep("none",2)), 6)
k <- rep(rep(c(0.1,Inf),3), 6)
pVac <- rep(seq(0, 1, 0.2), each = 6)


pars <- as.data.frame(cbind(Correlation, k, pVac))
pars$k <- as.numeric(k)
pars$pVac <- as.numeric(pVac)
```



```{r testmodel, eval = FALSE, include = FALSE}


vExpList <- list()


for(p in 1:nrow(pars)){
  
  correlation <- pars[p,1]
  k_transmission <- pars[p, 2]
  k_exposure <- pars[p,2]
  V_prop <- pars[p,3]
  
  print(paste0(p, " ", correlation, " ", k_transmission, " ", k_exposure))
  
  Runs_tracker <- 
    foreach (z = 1:runs, .combine = "rbind") %dopar% {
      state_update(
        ## Population Parameters ##
        state_tracker, # a vector of the the initial state of the population
        N,# population size (may exxclude in favour of calculating in state_update function)
        population_sample,# large or small population (determines sampling strategy)
        
        ## Model Parameters ##
        generations,# the number of generations (or timesteps) to run the model over
        output,# defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
        
        ## Epidemiological Parameters ##
        R_period, # the refractory period (number of timesteps that units are immune)
        R0_estimate, # average number of effective contacts per unit
        k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
        k_exposure, # dispersion parameter for exposure, required to generate probability distribution when independent of tranmsission
        heterogeneity, # population variation is “fixed” (constant over time), “variable” (changes each timestep) or “homogeneous” (all units have same potential)
        correlation, # relationship between transmission & exposure correlated postively (pos), negatively (neg) or unrelated (none)
        contact_list, # potential contacts of each unit, is NULL if pop_sample == "large"
        
        ## Reseeding Parameters ##
        reseed, # reseed 5 infections periodically during the first 10% of generations
        reseed_threshold, # threshold generation for reseeding events
        reseed_frequency,
        introductions, # number of new cases introduced when reseeding occurs
        
        ## Vaccination Parameters ##
        V_prop, # proportion to be vaccinated in each round
        V_start, # start generation for vaccination
        V_period, # duration of immunity from vaccination (by number of timesteps)
        V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
        V_schedule, # how frequent is vacciantion (ie every 2 "years")
        V_mode #random, target or non_target
      )
  
}

  vExpList[[p]] <- Runs_tracker

}


here()

save(vExpList, file = "ExposureExample1.RData")

```

```{r EndemicityAnalysis, include = FALSE}

load(here("data_080121/ExposureExample1.RData"))

newEnd <- sapply(vExpList, endemicity)

newEnddf <- as.data.frame(cbind(pars,"pEnd" = newEnd))



ggplot(data = newEnddf, aes(x = pVac, y = pEnd, group = factor(Correlation), fill = factor(Correlation), colour = factor(Correlation)))+
  geom_col(position = "dodge")+
  scale_x_continuous(breaks = seq(0,1,0.2))+
  scale_fill_discrete("Correlation")+
  scale_colour_discrete("Correlation")+
  xlab("Vaccine Coverage")+
  ylab("Proportion endemicity")+
  facet_wrap(~k)+
  theme_bw()


```
