---
title: "Vaccination and Heterogeneity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include= FALSE}
library(ggplot2)
library(here)
library(dplyr)
library(reshape2)
library(doParallel)
library(foreach)
library(readr)
#library(RColorBrewer)
#library(gridExtra) - for arranging plots side to side with grid.arrange
```

```{r parallelComputing, include = FALSE}

# Set Up for Parallel Computing

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


```


```{r functions, include = FALSE}

### Functions

# Load functions to run model:

source(here("Vfunctions/vStateUpdate_withHomogeneity.R")) # wrapper function, tracks epidemic in population over generations
source(here("Vfunctions", "V_DS_T2.R")) # updates state of population based on current generation
source(here("Vfunctions/vSI_withHomogeneity.R")) # simulation of infection with neg binom dist
source(here("Vfunctions", "effcIdGenT2.R")) # generating effective contact IDs
source(here("Vfunctions", "effcChecker.R")) # verifying effective contact IDs
source(here("Vfunctions", "reseedingT2.R")) # reseeding infection

```


```{r endStatsFn, include = FALSE}

#Include code for calculating median incidence for last 100 generations:

# Code to calculate median incidence at equilibrium (after 900Gens)

endStats <- function(mat){
  eq <- mat[, (ncol(mat)-100) :ncol(mat)] # subset last 100 generations of matrix
  id <- which(eq[,ncol(eq)] != 0) # identify which simulations become endemic (cases > 0 in final generation)
  if(length(id) == 0){ # if no simulations become endemic then all statistics set to 0
    M <- 0 
    iqr <- 0
    lq <- 0
    uq <- 0
    
  }else if( length(id) > 0 ){ # for simulations which become endemic...
    mEnd <- as.vector(eq[id,]) # subset last 100 generations of endemic simulations and format as vector
    # take the median, iqr, lower and upper quartiles over the last 100 generations for all simulations which become endemic 
    M <- median(mEnd) 
    iqr <- IQR(mEnd)
    lq <- quantile(mEnd)[2] # lower quartile
    uq <- quantile(mEnd)[4] # upper quartile
  }
  sumStat <- c(Med = M, iqr = iqr, lq = lq, uq = uq) # concatenate statistics
  return(sumStat)  # return vector of summary stats
  }
```

```{r pEndemicity, include = FALSE}
#Include code for calculating the proportion of simulations which become endemic for a given parameter set.


# Calculate proportion of simulations which become endemic

endemicity <- function(Tmat){ # input function is matrix of numbers infected across generations (columns) for multiple simulations (rows)
  
  P_endemic <- sum(Tmat[,ncol(Tmat)]!=0) / nrow(Tmat) # sum the number of simulations which result in 0 infections in the final generation (extinction), divided by the total number of simulaitons (number of row of matrix)

  return(P_endemic) # return the proportion of simulations which become endemic
}

```


### How does heterogeneity in the population impact vaccine effectiveness?

#### Random Vaccination

* How effective is random vaccination at preventing endemicity in populations with varying degrees of heterogeneity (in transmission)? 
* Include a homogeneous population for comparison
* Random Vaccination = individual-based control
  + Vaccinated units are selected at random from the population
  + The exposure and transmission potential of vaccinated units is set to 0, all other units remain the same.
  
##### Vaccination of Heterogeneous Populations: from Lloyd-Smith *et al.* (2005)

* Individual-specific measures outperform population-wide control measures
  + Population-wide measures (e.g. social distancing) are measures which reduce the transmission potential of every unit in the population by a set factor, c. 
  $v_c = (1-c)*v$
  
  + Individual-specific measures (e.g. vaccination) are interventions in which a proportion, c, of individuals have their transmission potential reduced to 0 
  $v_c = 0$ for vaccinated individuals
  $v_c = v$ for remainder of population
  
* Individual-specific measures increase outbreak heterogeneity, whilst population-wide measures decrease heterogeneity. A higher degree of heterogeneity promotes stochastic extinction, as shown previously.
* For both Individual-Specific, and Population-Wide control the threshold control effort for guaranteed extinction is: $c ≥ 1 - 1/R0$
  
### Model Parameters

* Choose parameter scenarios which result in endemicity for all values of k, in the absence of vaccination: 
  + $R0 = 3$: Intermediate R0 value, with Herd Immunity: $1-1/R0 = 0.66$
  + $Rperiod = 1$: No long-lasting immunity, promotes endemicity
  
* Model Conditions:
  + Generations = 2000 : sufficient generations for model to reach equilibrium and long enough for vaccination to take effect
  + Reseeding = TRUE : preventing early stochastic extinction
  
* Vaccination parameters:
  + Vaccination coverage = 0.6 - 0.7 : Values over herd immunity threshold (0.66)
  + Vaccination strategy = Random : vaccinated units are chosen randomly
  + Vaccination start generation = 1500 : after equilibrium is reached, with 500 generations until the end of the model
  + Vaccination period = Infinity : Vacciantion lasts for the duration of the model

```{r modelPars, eval = FALSE, include = FALSE}

## Model Definition ##
runs <- 500
generations <- 2000 # the number of generations (or timesteps) to run the model over
output <- "I_counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- 1 # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- Inf # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?

if(population_sample == "small"){
  contact_list1 <- replicate(N, c(1:N), simplify = FALSE)
  contact_list <- lapply(1:length(contact_list1), function(x) contact_list1[[x]][-x])
}else{
  contact_list <- vector()
}


## Reseeding Parameters ##
reseed <- TRUE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 100 # threshold generation for reseeding events
reseed_frequency <- 10
introductions <- 10 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 1500 # start generation for vaccination
V_period <- Inf # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 1 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed

```

```{r runModel, eval = FALSE, include = FALSE}

# With 2000 generations ##
het <- rep(c("fixed", "fixed", "fixed", "homogeneous"), 11)
k <- rep(c(0.1,0.3,Inf,Inf), 11)
pVac <- rep(seq(0.6, 0.7, 0.01), each = 4)

pars <- data.frame("k" = k, 
                   "het" = het, 
                   "pVac" = pVac)

R0_estimate <- 3
R_period <- 1
generations <- 2000
runs <- 500
V_start <- 1500

## Model ##

Vgen2000 <- list()

system.time(
  
  for (i in 1:nrow(pars)) {
  
  k_transmission <- pars[i, 1]
  heterogeneity <- pars[i, 2]
  V_prop <- pars[i, 3]
  
  print(i)
  
  Runs_tracker <-
    foreach (z = 1:runs, .combine = "rbind") %dopar% {
      state_update(
        state_tracker,# a vector of the the initial state of the population
        N, # population size (may exxclude in favour of calculating in state_update function)
        population_sample,# large or small population (determines sampling strategy)
        
        generations,# the number of generations (or timesteps) to run the model over
        output,# defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
        
        R_period, # the refractory period (number of timesteps that units are immune)
        R0_estimate, # average number of effective contacts per unit
        k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
        heterogeneity, # is heterogeneity in effective contacts "fixed" or "variable"
        contact_list, # potential contacts of each unit
        
        reseed, # reseed 5 infections periodically during the first 10% of generations
        reseed_threshold, # threshold generation for reseeding events
        reseed_frequency,
        introductions, # number of new cases introduced when reseeding occurs
        
        V_prop, # proportion to be vaccinated in each round
        V_start, # start generation for vaccination
        V_period, # duration of immunity from vaccination (by number of timesteps)
        V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
        V_schedule, # how frequent is vacciantion (ie every 2 "years")
        V_mode #random, target or non_target
      )
      
    }
  
  
  Vgen2000[[i]] <- Runs_tracker
  
}
)

save(Vgen2000, file = "Vgen2000_120221")

```


```{r vacPlot, echo = FALSE}
load(here("data_080121/Vgen2000_120221.RData"))

het <- rep(c("fixed", "fixed", "fixed", "homogeneous"), 11)
k <- rep(c(0.1,0.3,Inf,Inf), 11)
pVac <- rep(seq(0.6, 0.7, 0.01), each = 4)

pars <- data.frame("k" = k, 
                   "het" = het, 
                   "pVac" = pVac)


newEnd <- sapply(Vgen2000, endemicity)

newEnddf <- as.data.frame(cbind(pars,"pEnd" = newEnd))
newEnddf$condition <- rep(c("0.1","0.3", "Inf", "Homogeneous"))
newEnddf$condition <- factor(newEnddf$condition, levels = c("0.1", "0.3", "Inf", "Homogeneous"))


ggplot(data = newEnddf, aes(x = pVac, y = pEnd, group = factor(condition), fill = factor(condition), colour = factor(condition)))+
  geom_col(position = "dodge")+
  scale_x_continuous(breaks = seq(0.6,0.7,0.01))+
  scale_fill_discrete("k Value")+
  scale_colour_discrete("k Value")+
  xlab("Vaccine Coverage")+
  ylab("Proportion endemicity (500 simulations)")+
  theme_bw()

```

**Explanation** 

* With increasing heterogeneity, extinction occurs at lower vaccination coverage levels. 
* Useful metrics: 
  + What is the incidence of the endemic situations?
  + What do epidemic trajectories look like, for different k values?
  + If run for many simulations, are there any scenarios where HIT does not guarantee extinction?


### Individual Epidemic Trajectories

```{r trajectories, eval = FALSE, include = FALSE}
# condition 1: heterogeneity at 0.1,0.3,Inf

# condition <- rep(c("fixed", "fixed", "fixed", "homogeneous"), 11)
# k_transmission <- rep(c(0.1,0.3,Inf,Inf), 11)
# pVac <- rep(seq(0.6, 0.7, 0.01), each = 4)
het <- rep(c("fixed", "fixed", "fixed", "homogeneous"), 11)
k <- rep(c(0.1,0.3,Inf,Inf), 11)
pVac <- rep(seq(0.6, 0.7, 0.01), each = 4)

pars <- data.frame("k" = k, 
              "het" = het, 
              "pVac" = pVac)

R0_estimate <- 3
R_period <- 1
runs <- 1
output <- "counts"

trajectories <- list()


system.time(
  
  for (i in 1:nrow(pars)) {
  
  k_transmission <- pars[i, 1]
  heterogeneity <- pars[i, 2]
  V_prop <- pars[i, 3]
  
  print(i)
  
  Runs_tracker <-
    foreach (z = 1:runs, .combine = "rbind") %dopar% {
      state_update(
        state_tracker,# a vector of the the initial state of the population
        N, # population size (may exxclude in favour of calculating in state_update function)
        population_sample,# large or small population (determines sampling strategy)
        
        generations,# the number of generations (or timesteps) to run the model over
        output,# defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
        
        R_period, # the refractory period (number of timesteps that units are immune)
        R0_estimate, # average number of effective contacts per unit
        k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
        heterogeneity, # is heterogeneity in effective contacts "fixed" or "variable"
        contact_list, # potential contacts of each unit
        
        reseed, # reseed 5 infections periodically during the first 10% of generations
        reseed_threshold, # threshold generation for reseeding events
        reseed_frequency,
        introductions, # number of new cases introduced when reseeding occurs
        
        V_prop, # proportion to be vaccinated in each round
        V_start, # start generation for vaccination
        V_period, # duration of immunity from vaccination (by number of timesteps)
        V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
        V_schedule, # how frequent is vacciantion (ie every 2 "years")
        V_mode #random, target or non_target
      )
      
    }
  
  
  trajectories[[i]] <- Runs_tracker
  
}
)

save(trajectories, file = "vTrajectories120221.RData")
```

```{r trajanalysis, include = FALSE}

load(here("data_080121/vTrajectories120221.RData"))

trajectories_df <- as.data.frame(do.call("rbind", trajectories))
trajectories_df$states <- rep(c("S","I","R","V"), 44)
trajectories_df$cond <- rep(c(rep("0.1",4), rep("0.3",4), rep("Inf",4), rep("Homogeneous",4)), 11)
trajectories_df$cond <- factor(trajectories_df$cond, levels = c("0.1", "0.3", "Inf", "Homogeneous"))
trajectories_df$pVac <- rep(pVac, each = 4)

trajectoriesLong <- melt(trajectories_df, id.vars = c("states","cond","pVac"))
```


```{r trajectoriesPLots, echo = FALSE}

trajectoriesLong %>% 
  
  filter(
    pVac >=0.64, 
    pVac<=0.68, 
    states != "R") %>%
  
  ggplot(aes(x = variable, y = value, group = states))+
  geom_line(aes(colour = states))+
  facet_grid(pVac~cond)+
  scale_x_discrete(breaks = seq(800,1000,100), labels = as.character(seq(800,1000,100)))+
  coord_cartesian(xlim = c(800,1000))+
  xlab("Generations")+
  ylab("Incidence")+
  theme_bw()


trajectoriesLong %>% 
  
  filter(
    pVac == "0.66",
    states != "R") %>%
  
  ggplot(aes(x = variable, y = value, group = states))+
  geom_line(aes(colour = states))+
  facet_wrap(~cond)+
  scale_x_discrete(breaks = seq(800,1000,100), labels = as.character(seq(800,1000,100)))+
  coord_cartesian(xlim = c(800,1000))+
  xlab("Generation")+
  ylab("Incidence")+
  theme_bw()


```

**Explanation**
* Increased variation around median incidence for lower values of k, increases likelihood of stochastic extinction. 

