---
title: "Vaccine Simulations 020221"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include= FALSE}
library(ggplot2)
library(here)
library(dplyr)
library(reshape2)
library(doParallel)
library(foreach)
library(readr)
#library(RColorBrewer)
#library(gridExtra) - for arranging plots side to side with grid.arrange
```

```{r parallelComputing, include = FALSE}

# Set Up for Parallel Computing

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


```


### Functions

Load functions to run model:

```{r functions}
source(here("Vfunctions", "V_stateUpdateT2.R")) # wrapper function, tracks epidemic in population over generations
source(here("Vfunctions", "V_DS_T2.R")) # updates state of population based on current generation
source(here("Vfunctions", "S_to_I_T2.R")) # simulation of infection with neg binom dist
source(here("Vfunctions", "effcIdGenT2.R")) # generating effective contact IDs
source(here("Vfunctions", "effcChecker.R")) # verifying effective contact IDs
source(here("Vfunctions", "reseedingT2.R")) # reseeding infection

```


Include code for calculating median incidence for last 100 generations:

```{r endStatsFn}

# Code to calculate median incidence at equilibrium (after 900Gens)

endStats <- function(mat){
  eq <- mat[, (ncol(mat)-100) :ncol(mat)] # subset last 100 generations of matrix
  id <- which(eq[,ncol(eq)] != 0) # identify which simulations become endemic (cases > 0 in final generation)
  if(length(id) == 0){ # if no simulations become endemic then all statistics set to 0
    M <- 0 
    iqr <- 0
    lq <- 0
    uq <- 0
    
  }else if( length(id) > 0 ){ # for simulations which become endemic...
    mEnd <- as.vector(eq[id,]) # subset last 100 generations of endemic simulations and format as vector
    # take the median, iqr, lower and upper quartiles over the last 100 generations for all simulations which become endemic 
    M <- median(mEnd) 
    iqr <- IQR(mEnd)
    lq <- quantile(mEnd)[2] # lower quartile
    uq <- quantile(mEnd)[4] # upper quartile
  }
  sumStat <- c(Med = M, iqr = iqr, lq = lq, uq = uq) # concatenate statistics
  return(sumStat)  # return vector of summary stats
  }
```

Include code for calculating the proportion of simulations which become endemic for a given parameter set.
```{r pEndemicity}

# Calculate proportion of simulations which become endemic

endemicity <- function(Tmat){ # input function is matrix of numbers infected across generations (columns) for multiple simulations (rows)
  
  P_endemic <- sum(Tmat[,ncol(Tmat)]!=0) / nrow(Tmat) # sum the number of simulations which result in 0 infections in the final generation (extinction), divided by the total number of simulaitons (number of row of matrix)

  return(P_endemic) # return the proportion of simulations which become endemic
}

```


### Model Parameters

*Note: Running model for 3rd time to check output with vaccination - on 020221*
* See plot: Vplot_500Sims (with Rp = 1, R0 = 3, heterogeneity = fixed)

* Vaccination period = Infinity
* Vaccination strategy = Random
* Vaccination coverage between 0.5 - 0.75
* Refractory Periods of 1 (all endemicity)
* heterogeneity is fixed / variable
* reseeding = true


```{r pars1, eval = FALSE}

## Model Definition ##
runs <- 100
generations <- 1000 # the number of generations (or timesteps) to run the model over
output <- "I_counts" # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)

## Population Parameters ##

N <- 10000 # population size (may exxclude in favour of calculating in state_update function)
I0 <- 10
state_tracker <- c(rep("I", I0), rep("S", N - I0)) # initial disease state of units in population
population_sample <- "large" # large or small population (determines sampling strategy)

## Epidemiological Parameters ##

R_period <- 1 # the refractory period: 1 = no immunity, Inf = lifelong immunity
R0_estimate <- 3 # average number of effective contacts per unit
k_transmission <- Inf # overdispersion parameter, level of heterogeneity in effective contact potential: 0.1 = high heterogneiety, Inf = homogeneous

heterogeneity <- "fixed" # "fixed" or "variable", is unit transmission potential constant through time?
contact_list1 <- replicate(N, c(1:N), simplify = FALSE) 
contact_list <- lapply(1:length(contact_list1), 
                       function(x) contact_list1[[x]][-x]) 

## Reseeding Parameters ##
reseed <- TRUE # reseed 5 infections periodically during the first 10% of generations
reseed_threshold <- 100 # threshold generation for reseeding events
reseed_frequency <- 10
introductions <- 10 # number of new cases introduced when reseeding occurs

## Vaccination Parameters ##
V_prop <- 0 # proportion to be vaccinated in each round
V_start <- 900 # start generation for vaccination
V_period <- Inf # duration of immunity from vaccination (by number of timesteps)
V_rounds <- 1 # number of rounds (-1 for starting round) of vaccination in programme
V_schedule <- 0 # how frequent is vacciantion (ie every 2 "years")
V_mode <- "random" #random, target or non_target, NB: target/non-target only possible if heterogneiety is fixed

```

#### Run Model:

On 050221 run with the following parameters; 

* Rp_vals <- c(3,5,7)
* R0_vals <- c(2,3)
* k_vals <- c(0.1, 0.3, 0.5, Inf)
* pVac <- seq(0.5, 0.75, 0.05)
* heterogeneity <- "fixed"

```{r runModel, eval = FALSE}
Rp_vals <- c(3,5,7)
R0_vals <- c(2,3)
k_vals <- c(0.1, 0.3, 0.5, Inf)
pVac <- seq(0.5, 0.75, 0.05)
heterogeneity <- "fixed"
pars <- as.data.frame(expand.grid("R0" = R0_vals, 
                                  "R_period" = Rp_vals, 
                                  "k" = k_vals, 
                                  "pVac" = pVac
                                  ))

randVList <- list()

system.time(

for(i in 1:nrow(pars)){
    
    R0_estimate <- pars[i,1]
    R_period <- pars[i,2]
    k_transmission <- pars[i, 3]
    V_prop <- pars[i, 4]
    
    print(i)
    
    Runs_tracker <-
      foreach (z = 1:runs, .combine = "rbind") %dopar% {
       state_update(
         
          state_tracker, # a vector of the the initial state of the population
          N, # population size (may exxclude in favour of calculating in state_update function)
          population_sample, # large or small population (determines sampling strategy)
          
          generations, # the number of generations (or timesteps) to run the model over
          output, # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
          
          R_period, # the refractory period (number of timesteps that units are immune)
          R0_estimate, # average number of effective contacts per unit
          k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
          heterogeneity, # is heterogeneity in effective contacts "fixed" or "variable"
          contact_list, # potential contacts of each unit 
          
          reseed, # reseed 5 infections periodically during the first 10% of generations
          reseed_threshold, # threshold generation for reseeding events
          reseed_frequency,
          introductions, # number of new cases introduced when reseeding occurs
          
          V_prop, # proportion to be vaccinated in each round
          V_start, # start generation for vaccination
          V_period, # duration of immunity from vaccination (by number of timesteps)
          V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
          V_schedule, # how frequent is vacciantion (ie every 2 "years")
          V_mode #random, target or non_target
        )
        
      }
    
    
    randVList[[i]] <- Runs_tracker
    
}

)

#randVList - looking at random vaccination strategy with k = 0.1 and Inf and vac proportions of 0.5 - 0.75 (HIT = 0.66 with R0 =3)

```


```{r analysis}

load(here("data_080121/vList3.RData")) ## With Parameters below

# Vaccination parameters: 
p_vac <- seq(0.5, 0.75, 0.05) # proportion to be vaccinated in each round (try 0.6, 0.7, 0.8)
k_estimate <- c(0.1, Inf)
V_mode <- c("random", "target", "non_target")
pars <- expand.grid("k" =k_estimate, "mode" =V_mode, "pVac" =p_vac)


pEnd <- sapply(vList, endemicity)
  
End_df <- cbind(pars,"pEnd" = pEnd)
  
ggplot(data = End_df, aes(x = pVac, y = pEnd, group = factor(k), fill = factor(k), colour = factor(k)))+
  geom_col(position = "dodge")+
  scale_x_continuous(breaks = seq(0.5,0.75,0.05))+
  scale_fill_discrete("k Value")+
  scale_colour_discrete("k Value")+
  xlab("Vaccine Coverage")+
  ylab("Proportion endemicity (500 simulations)")+
  facet_wrap(~mode)+
  theme_bw()
  
  

```

```{r}
# Vaccination parameters: 



load("~/OneDrive - Royal Veterinary College/PhD_year1/Transmission Heterogeneity Model Jul-Aug 2020/Jan20_testing/vaccine_testing/vList050221.RData")
Rp_vals <- c(3,5,7)
R0_vals <- c(2,3)
k_vals <- c(0.1, 0.3, 0.5, Inf)
pVac <- seq(0.5, 0.75, 0.05)
pars <- as.data.frame(expand.grid("R0" = R0_vals, 
                                  "R_period" = Rp_vals, 
                                  "k" = k_vals, 
                                  "pVac" = pVac
))



pEnd <- sapply(randVList, endemicity)

End_df <- cbind(pars,"pEnd" = pEnd)

ggplot(data = End_df, aes(x = pVac, y = pEnd, group = factor(k), fill = factor(k), colour = factor(k)))+
  geom_col(position = "dodge")+
  scale_x_continuous(breaks = seq(0.5,0.75,0.05))+
  scale_fill_discrete("k Value")+
  scale_colour_discrete("k Value")+
  xlab("Vaccine Coverage")+
  ylab("Proportion endemicity (500 simulations)")+
  facet_grid(R_period~R0, labeller=label_both)+
  theme_bw()


```

```{r variableHet, eval = FALSE}

runs <- 100
R_period <- 3
R0_estimate <- 3
k_vals <- c(0.1, 0.3, 0.5, Inf)
pVac <- seq(0.5, 0.75, 0.05)
heterogeneity <- "variable"
pars <- as.data.frame(expand.grid("k" = k_vals, 
                                  "pVac" = pVac
                                  ))

varHList <- list()

system.time(

for(i in 1:nrow(pars)){
    
    k_transmission <- pars[i, 1]
    V_prop <- pars[i, 2]
    
    print(i)
    
    Runs_tracker <-
      foreach (z = 1:runs, .combine = "rbind") %dopar% {
       state_update(
         
          state_tracker, # a vector of the the initial state of the population
          N, # population size (may exxclude in favour of calculating in state_update function)
          population_sample, # large or small population (determines sampling strategy)
          
          generations, # the number of generations (or timesteps) to run the model over
          output, # defines the output as "matrix" (full disease state and ID of units) or "counts" (number of units in S,I,R states or "I_counts" (returns only count of infected units over time)
          
          R_period, # the refractory period (number of timesteps that units are immune)
          R0_estimate, # average number of effective contacts per unit
          k_transmission, # overdispersion parameter, level of heterogeneity in effective contact potential
          heterogeneity, # is heterogeneity in effective contacts "fixed" or "variable"
          contact_list, # potential contacts of each unit 
          
          reseed, # reseed 5 infections periodically during the first 10% of generations
          reseed_threshold, # threshold generation for reseeding events
          reseed_frequency,
          introductions, # number of new cases introduced when reseeding occurs
          
          V_prop, # proportion to be vaccinated in each round
          V_start, # start generation for vaccination
          V_period, # duration of immunity from vaccination (by number of timesteps)
          V_rounds, # number of rounds (-1 for starting round) of vaccination in programme
          V_schedule, # how frequent is vacciantion (ie every 2 "years")
          V_mode #random, target or non_target
        )
        
      }
    
    
    varHList[[i]] <- Runs_tracker
    
}

)



pEndvarH <- sapply(varHList, endemicity)

EndvarH_df <- cbind(pars,"pEnd" = pEndvarH)
library(ggplot2)
ggplot(data = EndvarH_df, aes(x = pVac, y = pEnd, group = factor(k), fill = factor(k), colour = factor(k)))+
  geom_col(position = "dodge")+
  scale_x_continuous(breaks = seq(0.5,0.75,0.05))+
  scale_fill_discrete("k Value")+
  scale_colour_discrete("k Value")+
  xlab("Vaccine Coverage")+
  ylab("Proportion endemicity (500 simulations)")+
  #facet_grid(R_period~R0, labeller=label_both)+
  theme_bw()

```
